---
output:
  md_document:
    variant: gfm
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include=FALSE}
options(width=80)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-")
```

# **multiLASSO: Directed Connectivity Search Using Regularized Regression**

The multiLASSO algorithm is a continually maintained R package.

Program developers are invited to submit changes here at the GitHub repository. 

# **The Basics**

+ Missing data is not a problem. 

+ Heterogeneous data is not a problem: 

  + No "group" or "common" network map will be forced unless it truly describes the majority. 
  
  + Individual-level nuances will surface after a group or common network map is derived (provided one exists).
  
  + If desired, subgroups of individuals with similar patterns of effects will be generated to aid the researcher in finding similar patterns among the varied individual models. 
  
+ Supports a large number of observed variables (capable of estimating a whole-brain network map). 
  + **NOTE:** Large numbers of variables will significantly increase run time.

  + Alternatively can support interaction effects between endogenous variables, or between exogenous and endogenous variables.

+ Can be freely downloaded by installing the package "multiLASSO" in R.

# **Running multiLASSO**

__1. Create two new folders (i.e., directories)__

+ Create a source folder for your time series. This can be anywhere that you have permission to read and write. 

+ Nothing can be in the source folder other than the time series data.

+ Create an output folder for your results. This must be different from the above folder. Alternatively, specifying an output folder will cause one to be created if permissions allow.

__2. Extract the time series for your variables__

  + Have each variable be a column, with the rows being the observation (e.g., scan in fMRI or a day in daily diary studies).

  + Substitute NA for missing values.

  + Have a separate file for each individual/session.

  + Put each file in the source folder you created in step 1. Do not put anything else in this folder.

  + Files must be either comma, space, or tab delimited.

__3. Installing multiLASSO with R__

  + Open an R script and enter into the console: `install.packages("multiLASSO")`
  
  + Once multiLASSO has been installed, you will need to load the package by entering: `library(multiLASSO)`
  
__4. Running multiLASSO__

  The *multiLASSO* function requires that you input:
  
  + The path to the directory containing your data.
  
  + How data are separated (e.g., comma-separated values).
  
  + Whether the data files contain a header row. 
  
  All other fields are optional and will go to defaults if no user input is provided. If no output directory is indicated, all information is stored as R objects (see tutorial linked above for details).
```{r, eval = FALSE}
output <- multiLASSO(
  data = '',          # source directory where your data are 
  out = NULL,         # output directory where you'd like your output to go (if NULL, output will only be saved in a list object)
  sep = NULL,         # how data are separated. "" for space; "," for comma, "/t" for tab-delimited
  header = TRUE,      # TRUE or FALSE, is there a header
  ar = TRUE,          # TRUE (default) or FALSE, start with autoregressive paths open
  plot = TRUE,        # TRUE (default) or FALSE, generate plots
  subgroup = FALSE,   # TRUE or FALSE (default), cluster individuals based on similarities in effects
  alpha = .5,         # option to control the elasticnet mixing parameter; alpha = .5 (default), alpha = 1 is the lasso penalty, alpha = 0 is the ridge regression penalty
  model_crit = 'bic', # model critera to use in model selection (default to BIC); other options include 'cv' (cross-validation), 'aic', 'aicc', 'hqc'
  penalties = NULL,   # option to specify a matrix of shrinkage parameters that will control the initial search for a group-level network map
  groupcutoff = .75,  # the proportion that is considered the majority at the group level
  subcutoff = .5      # the proportion that is considered the majority at the subgroup level
)        
```

While *multiLASSO* is running you will see information iterate in the command window. The algorithm will alert you when it is finished.

# **Output**

+ The output directory will contain:

  + __function_summary__: Contains details of function arguments and variable names (grouped by variable type) used in the current run.
  
  + __indivPathEstimates__: Contains details (iv & dv), regression estimate, and level (e.g., group, individual) for each path and each individual.
  
  + __groupPathCountsMatrix__: Contains counts of total number of paths, including contemporaneous, lagged, and interactions, estimated for the sample. The row variable is the predictor and the column variable is the outcome variable.
  
  + __groupPathCountsPresent__: Contains matrix of group path identities (1 = group-level path, 0 = other). The row variable is the predictor and the column variable is the outcome variable.
  
  + __groupMainEffectsPlots__: Produced if plot = TRUE. Contains figure with group, subgroup (if subgroup = TRUE), and individual-level paths for the sample. Black paths are group-level, green paths are subgroup-level, and grey paths are individual-level, where the thickness of the line represents the count. Contemporaneous paths are solid and lagged paths are dashed.
  
  + __groupInteractionsPlots__: Produced if plot = TRUE. Contains figure with group, subgroup (if subgroup = TRUE), and individual-level interaction paths for the sample. Black paths are group-level, green paths are subgroup-level, and grey paths are individual-level, where the thickness of the line represents the count. Interactions are represented by a knot combining the relevant predictors, with an arrow from the knot indicating the relevant outcome variable. Will only be created if interactions are specified.
  
+ In individual output directory (*where id represents the original file name for each individual*):

  + __*id*Betas__: Contains individual-level estimates of each path for each individual. Rows indicte predictors, columns indicate outcomes.
  
  + __*id*MainEffectsPlot__: Contains individual-level plots for main effects paths. Red paths represent positive weights and blue paths represent negative weights. Contemporaneous paths are solid and lagged paths are dashed.
  
  + __*id*InteractionsPlot__: Contains individual-level plots for interaction paths. Interactions are represented by a knot combining the relevant predictors, with an arrow from the knot indicating the relevant outcome variable. Red paths represent positive weights and blue paths represent negative weights. Contemporaneous paths are solid and lagged paths are dashed. Will only be created if interactions are specified.
  
# **FAQ**

**How many time points do I need?**
This is a difficult question since it will be related to the number of variables you are using. Rules of thumb for any analysis can generally be used: the more the better! Having at least 100 time points is recommended, but adequate results have been obtained in simulation studies with only T = 60.

**Do all individuals have to have the same number of observations (T)?**
No.

**How many people do I need in my sample?**
For *multiLASSO*, reliable results are obtained with as few as 10 participants. Remember that in this context, power to detect effects is determined by the number of time points rather than the number of individuals. Still, having at least 10 individuals helps *multiLASSO* to detect signal from noise by looking for effects that consistently occur.

**What do I do if I obtain an error?**
Do some initial trouble-shooting. 

1. Ensure that all of your individuals have the same number of variables (columns) in their data sets.

2. Ensure that all variables have variability (i.e., are not constant). *multiLASSO* will let you know if this is the case.

3. Ensure your path directories are correct.

4. Ensure that the columns are variables and the rows contain the observations across time.

5. If all of this is correct, please email the error you received, code used to run *multiLASSO*, and the data (we promise not to use it or share it) to: gimme@unc.edu.